Determining the alternatives for scalar implicature

### 2016-01-29 Paper/Paper.tex
fixed plotting for literal listener and cut out part of model results (now it's in the general discussion)

Experiment 1 was conducted to approximate literal listener semantic
distributions \(p_{L_0}(m \mid u)\) for five pairs of scalar items taken
from Tiel (2014). We include three conditions in Experiment 1,
corresponding to the sets of alternatives within a scale that
participants were presented with: two alternatives (``entailment''),
four alternatives, and five alternatives. The two-alternative entailment
condition makes a test of the hypothesis that the two members of the
classic Horn scale (Horn, 1972) are the only alternatives necessary to
predict the strength of listeners' pragmatic inference. The four- and
five-alternative conditions then add successively more alternatives to
test whether including a larger number of alternatives will increase
model
fit.\footnote{Note that alternatives in the four- and five-alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}

@@

To test whether there were between experiment differences in
acceptability judgments, we ran a mixed effects model, regressing
acceptability judgements on scale, number of stars and experiment type,
with by subject and by word random effects. Results indicate no
significant differences between Experiment 1a and 1c \(b =\)-0.05,
\(Z =\) -0.53, \(p =\) 0.59 or between 1b and 1c \(b =\) -0.04, \(Z =\)
-0.52 \(p =\) 0.6 and the addition of our experiment predictor did not
significantly improve model fit when compared to a model without the
experiment type variable using ANOVA \(\chi^2(\) 2 \() =\) 0.43, \(p =\)
0.81.

@@

Experiment 3 was conducted to measure pragmatic judgments
\(p_{L_1}(m \mid u)\) for five pairs of scalar items taken from Tiel
(2014). We include two conditions in Experiment 3. In the
two-alternative condition, participants made judgments for items
included in the entailment scales. In the four-alternative condition,
participants made judgments for the entialment items and also the top
two alternatives elicited for each scale in Experiment 2. Running two
pragmatic judgment conditions allowed us to rule out the potential
effects of having a larger set of alternatives during the pragmatic
judgment elicitation and also provided two sets of human judgments to
compare with model predictions.
\footnote{Note that alternatives in the four-alternative condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}

@@

To test whether there were between experiment differences in pragmatic
judgments, we ran a mixed effects model, regressing pragmatic judgements
on scale and experiment with subject and word level random effects.
Results indicate there were no significant differences between
Experiment 3a and 3b for \(b =\) 0.05, \(t(\) 150 \() =\) 1.04, \(p =\)
0.3 and the addition of our experiment predictor did not significantly
improve model fit when compared to a model without the experiment
variable using ANOVA \(\chi^2(\) 1 \() =\) 1.13, \(p =\) 0.29.

@@

Using literal semantic data from Experiments 1a,b,c we conducted three
simulations with our model. Each simulation used the specific literal
semantic data to specify the scale representation available to our
model. The ``entailment'' model used only the original pair of scalar
items pulled from van Tiel (2014). These data were measured in
Experiment 1a. The ``Top two'' model extended the original set of
alternatives with the top two alternatives generated in Experiment 2.
These data were measured in Experiment 3a. The ``Full'' model included
the full set of alternatives, including a neutral valenced alternative.
These data were measured in Experiment 4.

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
Model & Exp. 3a & Exp 3b \\ 
  \hline
Entailment & 0.54 & 0.57 \\ 
  Top\_two & 0.84 & 0.90 \\ 
  Full & 0.86 & 0.90 \\ 
   \hline
\end{tabular}
\caption{Model fit measured by correlation coefficients} 
\end{table}

@@

By varying the type of scale representations available to our Bayesian
model we investigated the effects of alternatives on scalar implicature.
Model fit with human judgement was significantly improved by the
inclusion of alternatives beyond the typical ``strong'' and ``weak''
scalar items. In fact, we found that both neutral and negative valence
scalar items contribute to human-like implicature generation within our
framework.

>>>>

Experiment 1 was conducted to approximate literal listener semantic
distributions \(p_{L_0}(m \mid u)\) for five pairs of scalar items taken
from Tiel (2014). We include three conditions in Experiment 1,
corresponding to the sets of alternatives within a scale that
participants were presented with: two alternatives (``entailment''),
four alternatives, and five alternatives. The two-alternative entailment
condition makes a test of the hypothesis that the two members of the
classic Horn scale (Horn, 1972) are the only alternatives necessary to
predict the strength of listeners' pragmatic inference. The four- and
five-alternative conditions then add successively more alternatives to
test whether including a larger number of alternatives will increase
model
fit.\footnote{Note that alternatives in the four- and five-alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiments 1b,c tests whether the total set of
alternatives queried during literal semantic elicitation effects
compatibility judgments. (If it does we should see differences in
compatibility judgments between experiments.) We address this possibiliy
in the results section.

@@

To test our secondary hypothesis, that the different sets of scalar
items queried might result in differences in compatibility judgments
between experiments, we ran a mixed effects model. We regressed
compatibility judgments on scale, number of stars and experiment
(1a,b,c), with subject and word level random effects, which was the
maximal structure that converged. Results indicate no significant
differences between Experiment 1a and 1c \(b =\)-0.05, \(Z =\) -0.53,
\(p =\) 0.59 or between 1b and 1c \(b =\) -0.04, \(Z =\) -0.52 \(p =\)
0.6 and the addition of our experiment predictor did not significantly
improve model fit when compared to a model without the experiment
variable using ANOVA \(\chi^2(\) 2 \() =\) 0.43, \(p =\) 0.81.

@@

Experiment 3 was conducted to measure pragmatic judgments
\(p_{L_1}(m \mid u)\) for five pairs of scalar items taken from Tiel
(2014). We include two conditions in Experiment 3. In the
two-alternative condition, participants made judgments for items
included in the entailment scales. In the four-alternative condition,
participants made judgments for the entialment items and also the top
two alternatives elicited for each scale in Experiment 2. Running two
pragmatic judgment conditions allowed us to rule out the potential
effects of having a larger set of alternatives during the pragmatic
judgment elicitation and also provided two sets of human judgments to
compare with model predictions.
\footnote{Note that alternatives in the four-alternative condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 3b tests whether the total set of
alternatives queried in Experiments 3a,b effects pragmatic judgments.
(If it does we should see differences in pragmatic judgments between
experiments.) We address this possibiliy in the results section.

@@

To test our secondary hypothesis, that the different sets of scalar
items queried might result in differences in pragmatic judgments between
experiments, we ran a mixed effects model. We regressed pragmatic
judgements on scale and experiment (3a,b) with subject and word level
random effects, which was the maximal structure that converged. Results
indicate there were no significant differences between Experiment 3a and
3b for \(b =\) 0.05, \(t(\) 150 \() =\) 1.04, \(p =\) 0.3 and the
addition of our experiment predictor did not significantly improve model
fit when compared to a model without the experiment variable using ANOVA
\(\chi^2(\) 1 \() =\) 1.13, \(p =\) 0.29.

@@

Using literal semantic data from Experiments 1a,b,c we conducted three
simulations with our model. Each simulation used the specific literal
semantic data to specify the scale representation available to our
model.

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
Model & $\alpha$ & Exp. 3a & Exp. 3b \\ 
  \hline
Two alts &   9 & 0.54 & 0.57 \\ 
  Two alts + generic negative  &   6 & 0.62 & 0.66 \\ 
  Four alts &   4 & 0.84 & 0.90 \\ 
  Five alts &   4 & 0.86 & 0.90 \\ 
   \hline
\end{tabular}
\caption{Model performance with fitted alpha levels. Model fit assessed through correlation with human judgments in our two Pragmatic listener experiments (3a,b)} 
\end{table}

{\centering \includegraphics{figs/fiveAltsScatter-1}

\caption[Model fit]{Model fit: model predictions are plotted on the x-axis. Human judgments from Exerpiment 3b are plotted on the y-axis. Coloration denotes star rating for which judgment was made.}\label{fig:fiveAltsScatter}
\end{figure*}
\end{CodeChunk}

@@

By varying the type of scale representations available to our Bayesian
model we investigated the effects of alternatives on scalar implicature.
Model fit with human judgments was significantly improved by the
inclusion of alternatives beyond the typical ``strong'' and ``weak''
scalar items. The ``Two alternative'' model contained only entailment
items, which, under classic accounts, should be sufficient to generate
implicature. Model performance with this limited alternative set was
poor when compared to human judgments from Experiment 3a,b (row 1, Table
\ref{xtable}). Building off findings in Franke (2014), in which the
best-fitting model assigned substantial weight to \emph{none} as an
alternative, we included a ``generic'' negative alternative that was
only compatible with 1-star. Model fit for this modified ``Two
alternative with generic negative'' alternative set did improve, however
overall correlation was still low (row 2, Table \ref{xtable}). Model fit
jumped substantially by adding the top-two alternatives for each scale
(row 3, Table \ref{xtable}). Finally, our ``Five alternative'' model
included a neutral alternative along with the original entailment pair
and top two (row 4, Table \ref{xtable}).

While model improvement seemed to be related to the number of
alternatives present, we did not see substantial differences between our
four- and five-alternative models. One possibility is that alternatives
are differentially salient in context. Our current framework makes the
simplifying assumption that all alternatives are equally salient.
However, it may well be the case that while both \emph{all} and
\emph{most} are \emph{present} during pragmatic enrichment of
\emph{some}, \emph{all} is more salient and therefore should carry
greater weight in our model. Findings from Experiment 2, in which
participants generated plausible alternatives, supports this intuition,
as some alternatives appeared more often than others. This distinction
between binary \emph{presence} and graded \emph{salience} is an
important one that deserves further investigation.

It is also likely that the precise set of alternatives present during
implicature are domain dependent. Our current empirical paradigm
elicited literal semantics, pragmatic judgments and plausible
alternatives all within the restricted domain of food reviews.
Participant data may have differed greatly if we had instead asked them
to respond in a movie review paradigm, which would have been acceptable
for a number of our stimuli (ie \emph{memorable}/\emph{unforgettable},
\emph{liked}/\emph{loved}, etc.).

Considering the possibility that alternative sets may be domain specific
even for scalar items may provide a way forward in uniting generalized
and particularized implicatures. Findings from our investigation
indicate that human literal semantic and pragmatic judgments are often
graded and diverse, much like the set of plausible alternatives that
give rise to pragmatic inferences.
